---
title: "Time series analysis, dataset 2"
output:
  html_document:
    df_print: paged
---

In this document it a time series dataset is going to be check to see if it is free from noise and it
can be reasonable describe by a closed mathematical formula.

# Install Libraries
```{r}
#install.packages('data.table')
#install.packages('TSA')
#install.packages('ggplot2')
#install.packages('forecast')
```


# Load dataset and perform visual inspection
```{r}
library(data.table)
data2=fread('data/data2.csv')
plot(seq(1:900),data2[1:900]$V1)
```

# Extract espectral distribuition
```{r}
library(TSA)
p=periodogram(data2$V1)
#plot(p)
```

# Sort and extract the top 20 frequencies
```{r}
wdata=data.table(freq=p$freq,spec=p$spec)
wdata[order(-spec)][1:20]
```
The lowest relevent period is `r round(1/wdata[order(-spec)][1:20][,max(freq)])`.

Let's smooth and explore the data in order visualyze.

```{r}
#w=2*pi/Period
train_length=as.integer(length(data2$V1)*3/4)
train=seq(1:train_length)
data_phase=data.table(x=seq(1:length(data2$V1)),y=data2$V1,type='test')
data_phase[train]$type='train'
library(forecast)
data_phase[,y_50:=ma(y,50)]
data_phase[,y_100:=ma(y,100)]
data_phase[,y_200:=ma(y,200)]
ggplot(data_phase)+geom_line(aes(x,y))+
  geom_line(aes(x,y_50),color='green')+
  geom_line(aes(x,y_100),color='blue')+
  geom_line(aes(x,y_200),color='red')
```


The plot shows that the fluctuations may be periodic on a small scale, and there is trends on a macro scale
that is hard to see any obvious pattern.
```{r}
total=dim(data_phase)[1]
max_windows_size=1000
max_lag=100
offset_size=(total-max_windows_size-max_lag)
lags=function(lag,windows_size){
  corrs=sapply(seq(1,offset_size,20),function(offsets){
    lagoffset=offsets+lag
    cor(data_phase[offsets:(offsets+windows_size)]$y,
             data_phase[lagoffset:(lagoffset+windows_size)]$y)
  })
  data.table(lag=lag,offset=(1:length(corrs)-1)*20+1,cor=corrs,window=windows_size)}
gen=function(windows_size){
  rbindlist(lapply(seq(1,max_lag,20),function(lag) lags(lag,windows_size)))}
corr=rbindlist(lapply(seq(10,max_windows_size,100),gen))
ggplot(corr,aes(x=offset,y=cor,color=factor(lag)))+geom_line()+facet_wrap(~window)
```

```{r}
fit_model=auto.arima(data_phase[train]$y)
predict_data= forecast(fit_model,length(data_phase[type=='test']))
data_phase$fit=c(fit_model$fitted.values,predict_data)
library(ggplot2)
ggplot(data=data_phase[type=='test'][1:50])+geom_line(aes(x=x,y=fit),color='blue')+geom_point(aes(x=x,y=y))
```
