---
title: "Time series analysis, dataset 1"
output:
  html_document:
    df_print: paged
---

In this document it a time series dataset is going to be check to see if it is free from noise and it
can be reasonable describe by a closed mathematical formula.

# Install Libraries
```{r}
#install.packages('data.table')
#install.packages('TSA')
#install.packages('ggplot2')
```


# Load dataset and perform visual inspection
```{r}
library(data.table)
data1=fread('data/data1.csv')
plot(seq(1:100),data1[1:100]$V1)
```

# Extract espectral distribuition
```{r}
library(TSA)
p=periodogram(data1$V1)
plot(p)
```

# Sort and extract the top 6 frequencies
```{r}
wdata=data.table(freq=p$freq,spec=p$spec)
wdata[order(-spec)][1:6]
```

# Convert the top frequency to Period.
```{r}
Period=round(1/wdata[order(-spec)][1]$freq)
Period
```

Now that it has been verified that a clear frequency dominates the behavior of the data, a sinusoidal fit
of the form: $$y = A Sin(\omega x + \phi)$$ is to be performed.
From trigonometric identity it can be seen that
$$y = a Sin(\omega x) + b Cos(\omega x)$$
$$ A = \sqrt{a^2+b^2}$$
The parameter can now be fitted by linear regression:
```{r}
w=2*pi/Period
train_length=as.integer(length(data1$V1)*3/4)
train=seq(1:train_length)
data_phase=data.table(x=seq(1:length(data1$V1)),y=data1$V1,type='test')
data_phase[train]$type='train'
fit_model = lm(y ~ sin(w*x)+cos(w*x),data=data_phase[train])
w
```

```{r}
summary(fit_model)
```

```{r}
library(ggplot2)

predicted=predict(fit_model,newdata=data_phase[-train])
fit=c(fit_model$fitted.values,predicted)
data_phase$fit=fit

data_phase[,residuals:=((y-fit))]
ggplot(data_phase,aes(x=fit,y=residuals,color=type,size=type))+geom_point()
```

The fit is essentially perfect, that is, the parameters incertainties are smaller than the available precision of the estimates: 
```{r}
parameters=coefficients(fit_model)
errs=summary(fit_model)$coefficients[5:6]
param=data.table(a=parameters[2],a_err=errs[1],
b=parameters[3],b_err=errs[2])
param[,A:=sqrt(a^2+b^2)]
param[,phi:=asin(a/A)]
```
$$y = Sin(0.06283185 x + 0.6336282)$$
```{r}
plot(data_phase[type=='test']$y,data_phase[type=='test']$fit)
```
```{r}
ggplot(data_phase[type=='test'])+geom_line(aes(x=x,y=y),color='red')+
geom_line(aes(x=x,y=sin(w*x+param$phi[1]+0.4)),pch=2)
```